\begin{acknowledgements}

\addchaptertocentry{\acknowledgementname}
Five years ago, when I decided to start studying to be part of the IT industry someday, 
I had no idea what was waiting for me. 
I did not get into computer science before starting university, and because of that 
I wanted to ask Prof. Rocco Oliveto - which at that time already was the chair of the bachelor program - for his advice 
in this regard. Even though I did not know why, on talking to him I immediately felt the path I was about to choose was right for me. 
I trusted him and I trusted my instincts. \\
Today, after five years - one of which I studied abroad - I think my place in the world is starting to become clear.
Never have I ever felt more confident on the decision I made. Yet I am a person who more often than needed struggles with self-confidence.\\
\newline
I owe Prof. Rocco Oliveto - my advisor and one of the most inspiring teachers I have encountered
in my course of study - a huge thank you, for everything mentioned above and more. Not only was he the reason why I started my journey in one of the most impactful and topic-broad field, 
but he also gave me a lot of opportunities to grow educationally and personally through both his academic lectures and life-lessons.
Thank you for supporting me during the last few years and for being part of the professional future I have just started to create.

I would also like to thank Prof. Rocco Oliveto, Prof. Gabriele Bavota, and all the UNIMOL and USI staff who gave me the life-changing opportunity to 
achieve the double degree. 
Spending the last year of my Masters at Università della Svizzera Italiana, was pivotal. Even though it was
really tough, it came at a moment in my life when I needed it. 
Getting through that extremely challenging and demanding 
time was genuinely essential in teaching to me what really matters, not only career-wise, but also in life. Thank you 
for the unique opportunity you gave me, and your wholehearted support.

I thank Prof. Gabriele Bavota who since the very first moment I met him, made himself available for any kind of support. 
Thank you for not missing any opportunity to remind me that there was someone on the other side who believed in me. Also, thank you 
for giving me the chance to work on the master thesis with your advice. I have learned so much.

This piece of work would not have been possible without the help of Antonio Mastropaolo. 
Thank you for all your time and your words of advice, I will not foget them.\\
\newline
I owe a huge thank you to Tania Clarke. Not only is she the best English teacher I have ever had, 
but for the past two years she has been a very important figure for me. A source of knowledge
of a language that I have always wanted to perfect and learn more about. Thank you for always being supportive.
Studying aside, your letting me talk and your simple and direct advice helped me.\\
\newline
There are no words to thank my family, source of unconditional love and endless support. 
The only thing in the whole world which I cannot live without;
the source of strength, courage, foresight, wisdom; the reason why I am who I am.
Thank you mom and dad for giving me the right discipline; thanks for being my pillars, 
empathic objectivity on one side and hopeful rationality on the other. Thank you Ugo and Vito, 
brothers as annoying as they are attentive and loving. Thank you for always being by my side. What would I do without you?\\
\newline
Special thanks go to Martina, Antonio, and Silvia. 
You could see how I really am and I truly feel extremely grateful 
for standing there for me, and not only when I really needed it. 
I will never forget your constant support, that has emerged a little more in the past year, but that was 
just a testament to a solid and sincere friendship. I am glad of having you on my side and I will always be by yours.
You can count on it!\\
\newline
My abroad experience would have not been the same if I had not met Ottavio, Daniel, Gianlorenzo, Federico, and Isabella. 
Thanks for all the sharing views we talked about, the fun we had together, and the memories we collected. I feel very grateful to have met you.
You now take up a part of my heart.\\
\newline
Thank you all.
\end{acknowledgements}